Optimal Task Phasing for End-To-End Latency in Harmonic Systems and Semi-Harmonic Automotive Systems

### 2024-10-02 paper.tex
comments with matthias

	In the remainder of this section, we consider the case that $\Tbb_E$ is not max-harmonic. 
	Specifically, this is achieved in the following two cases:
	\begin{enumerate}[label=(\roman*),start=4]
		\item $\Tmaxone = 5$ and $\Tmaxtwo=2$ \label{item:automotive_max_harm_case4}
		\item $\Tmaxone = 50$ and $\Tmaxtwo=20$ \label{item:automotive_max_harm_case5}
	\end{enumerate}
	In both cases, the hyperperiod $H = \lcm(T_{\tau_1}, \dots, T_{\tau_n})$ of tasks of $\Tbb_E$ is two times the highest period, i.e., $H = 2\cdot \Tmaxone$.
	This observation allows us to determine the end-to-end latency by constructing only two partitioned job chains.

@@

	Therefore, to minimize the end-to-end latency, we need to minimize the maximal length of both $\pc^p_{W_p}$ and $\pc^p_{W_p+1}$.
	However, Figure~\ref{fig:automotive_example} demonstrates that it is not possible to eliminate the slack between write- and read-events of subsequent jobs of both $\pc^p_{W_p}$ and $\pc^p_{W_p+1}$.
	%
	\begin{figure}
		\centering
		\subfloat[Minimize $\ell(\pc^p_{W_p})$.]{
		\begin{tikzpicture}[yscale=0.25, xscale=0.27]

@@

@@

			\end{scope}
			\end{tikzpicture}	
		\label{fig:automotive_example:optimal_first}}
		\\
		\subfloat[Minimize $\ell(\pc^p_{W_p+1})$.]{
			\begin{tikzpicture}[yscale=0.25, xscale=0.27]
			\begin{scope}[shift={(0,6)}] % task one
				\taskname{$\tau_1$}

@@

			\end{scope}
			\end{tikzpicture}	
		\label{fig:automotive_example:optimal_second}}
		\caption{Example of three tasks demonstrating that not both $\ell(\pc^p_{W_p})$ and $\ell(\pc^p_{W_p+1})$ can be minimized simultaneously. The partitioned job chain $\pc^p_{W_p}$ is marked in red and $\pc^p_{W_p+1}$ is marked in blue. The jobs belonging to both are half red and half blue. Gaps are depicted as left-right arrows.}
		\label{fig:automotive_example}
	\end{figure}
	%
	The example consists of a cause-effect chain $E = (\tau_1 \to \tau_2 \to \tau_3 )$ with three tasks.
	The tasks have periods $T_{\tau_1} = 5$, $T_{\tau_2} = 1$ and $T_{\tau_3} = 2$.
	Therefore, that cause-effect chains falls in case \ref{item:automotive_max_harm_case4}, and $\Tbb_E$ is not max-harmonic.
	If we minimize $\ell(\pc^p_{W_p})$, the write-event of $\tau_2$ and the read-event of $\tau_3$ do not align for $\pc^p_{W_p+1}$, as depicted in Figure~\ref{fig:automotive_example:optimal_first}.
	Conversely, if we minimize $\ell(\pc^p_{W_p+1})$, the write-event of $\tau_2$ and the read-event of $\tau_3$ do not align for $\pc^p_{W_p}$, as depicted in Figure~\ref{fig:automotive_example:optimal_second}.

@@

>>>>

	In the remainder of this section, we consider the scenario that $\Tbb_E$ is not max-harmonic. 
	Specifically, this is achieved in the following two cases:
	\begin{enumerate}[label=(\roman*),start=4]
		\item $\Tmaxone = 5$ and $\Tmaxtwo=2$ \label{item:automotive_max_harm_case4}
		\item $\Tmaxone = 50$ and $\Tmaxtwo=20$ \label{item:automotive_max_harm_case5}
	\end{enumerate}
	In both cases, the hyperperiod $H = \lcm(T_{\tau_1}, \dots, T_{\tau_n})$ of tasks of $\Tbb_E$ is two times the highest period, i.e., $H = 2\cdot \Tmaxone$.
	This observation allows us to determine the end-to-end latency by constructing only two partitioned job chains.

@@

	Building upon the results of the previous lemma, to minimize the end-to-end latency, we need to minimize the maximal length of both $\pc^p_{W_p}$ and $\pc^p_{W_p+1}$.
	However, while it is possible to achieve either $\pc^p_{W_p} = \sum_{i=1}^n T_{\tau_i} + \Tmaxone$ or $\pc^p_{W_p+1} = \sum_{i=1}^n T_{\tau_i} + \Tmaxone$, 
	Figure~\ref{fig:automotive_example} demonstrates that it is not possible to achieve both simultaneously.
	%
	\begin{figure}
		\centering
		\subfloat[Minimize $\ell(\pc^1_{1})$.]{
		\begin{tikzpicture}[yscale=0.25, xscale=0.27]

@@

				\timeline{0}{23}{}

@@

			\begin{scope}[shift={(0,0)}] % task four
				\taskname{$\tau_4$}

				\fill[color=red] (13,0) rectangle (14,1);

				\fill[color=blue] (19,0) rectangle (20,1);

				\exec{9}{10}
				\exec{10}{11}
				\exec{11}{12}
				\exec{12}{13}
				\exec{13}{14}
				\exec{14}{15}
				\exec{15}{16}
				\exec{16}{17}
				\exec{17}{18}
				\exec{18}{19}
				\exec{19}{20}
				\exec{20}{21}
				\exec{21}{22}
				\execopen{22}{22.5}

			\end{scope}
			\end{tikzpicture}	
		\label{fig:automotive_example:optimal_first}}
		\\
		\subfloat[Minimize $\ell(\pc^1_{2})$.]{
			\begin{tikzpicture}[yscale=0.25, xscale=0.27]
			\begin{scope}[shift={(0,6)}] % task one
				\taskname{$\tau_1$}

@@

			\end{scope}
			\begin{scope}[shift={(0,0)}] % task four
				\taskname{$\tau_4$}

				\fill[color=red] (14,0) rectangle (15,1);

				\fill[color=blue] (18,0) rectangle (19,1);

				\exec{8}{9}
				\exec{9}{10}
				\exec{10}{11}
				\exec{11}{12}
				\exec{12}{13}
				\exec{13}{14}
				\exec{14}{15}
				\exec{15}{16}
				\exec{16}{17}
				\exec{17}{18}
				\exec{18}{19}
				\exec{19}{20}
				\exec{20}{21}
				\exec{21}{22}
				\execopen{22}{22.5}

			\end{scope}
			\end{tikzpicture}	
		\label{fig:automotive_example:optimal_second}}
		\caption{Example of four tasks demonstrating that not both $\ell(\pc^1_{1})$ and $\ell(\pc^1_{2})$ can be minimized simultaneously. The partitioned job chain $\pc^1_{1}$ is marked in red and $\pc^1_{2}$ is marked in blue. The jobs belonging to both are half red and half blue. Gaps are depicted as left-right arrows.}
		\label{fig:automotive_example}
	\end{figure}
	%
	Instead, a small gap of one time unit has to be considered either for $\pc^p_{W_p}$ or for $\pc^p_{W_p+1}$.
	The example consists of a cause-effect chain $E = (\tau_1 \to \tau_2 \to \tau_3 \to\tau_4 )$ with four tasks.
	The tasks have periods $T_{\tau_1} = 5$, $T_{\tau_2} = 1$, $T_{\tau_3} = 2$ and $T_{\tau_4} = 1$.
	Therefore, that cause-effect chains falls in case \ref{item:automotive_max_harm_case4}, and $\Tbb_E$ is not max-harmonic.
	By definition, $p=1$ and $W_p=1$, and we consider the $1$-partitioned job chains 
	\begin{align}
		\pc^1_1 
		&= ((\tau_1(j_{1,1})),(\tau_1(j'_{1,1}) \to \dots \to \tau_4(j'_{1,4})))
		\\
		\pc^1_2 
		&= ((\tau_1(j_{2,1})),(\tau_1(j'_{2,1}) \to \dots \to \tau_4(j'_{2,4}))).
	\end{align}
	If we minimize $\ell(\pc^1_{1})$, the write-event of $\tau_2(j'_{2,2})$ and the read-event of $\tau_3(j'_{2,3})$ do not align for $\pc^1_{2}$, as depicted in Figure~\ref{fig:automotive_example:optimal_first}.
	Conversely, if we minimize $\ell(\pc^1_{2})$, the write-event of $\tau_2(j'_{1,2})$ and the read-event of $\tau_3(j'_{1,3})$ do not align for $\pc^1_{1}$, as depicted in Figure~\ref{fig:automotive_example:optimal_second}.
	In the following, we identify the cause for such gaps, and quantify their size.

	The reason for the gaps in Figure~\ref{fig:automotive_example} is that jobs of $\tau_1$ can only be multiples of $\Tmaxone$ apart, and jobs of $\tau_3$ can only be multiples of $\Tmaxtwo$ apart. 
	More specifically,

	More specifically, if we consider the two $1$-partitioned job chains

	then the jobs $\tau_1(j_{1,1})$ and $\tau_1(j_{2,1})$,

	\mario{We should lead this discussion towards
	\begin{equation}
		\lat(E) \geq \sum_{i=1}^n T_{\tau_i} + \Tmaxone + \min(\Tmaxone, \ceiling{\frac{\nu_E}{2}} \gamma_E)
	\end{equation}
	}

@@

- arbitrary periods: identify max-harmonic subchains, determine gap-size, try to distribute evenly. (Becomes too complex because then 3 or even more partitioned chains need to be considered for the distribution.)
